---
title: 'Coronavirus!'
author: "Ram Mukund Kripa"
output: md_document
toc: no
---

### Introduction

I am starting out with a very simple question: Are rich countries or poor countries affected more by the coronavirus?
I am using GDP as an indicator of "richness" even though this may or may not be the most accurate representation possible.


``` {r}
# packages and data is read in
library(dplyr)
library(ggplot2)
corona = read.csv("~/Downloads/owid-covid-data.csv")
recent = filter(corona,date=='2020-05-25')
recent = filter(recent,location!='World')
```

### Exploration

Are GDP and total cases correlated?

```{r}
# is there a correlation between gdp per capita and total cases?
ggplot(recent,aes(x=gdp_per_capita,y=total_cases))+
  geom_point(color="skyblue")+
  geom_smooth(method='lm')
x = lm(gdp_per_capita ~ total_cases,recent)
summary(x)
#seems to be a slightly positive correlation

```

This didn't really yield any conclusive results.
However, what if we compared the GDP per capita with the number of cases per million?
If we compare both on a scale of per person, we should get more reliable results!

```{r}
ggplot(recent,aes(x=gdp_per_capita,y=total_cases_per_million))+
  geom_point(color="skyblue")+
  geom_smooth(method='lm')
x = lm(gdp_per_capita ~ total_cases_per_million,recent)
summary(x)
```

From this it would seem that higher GDP indicates more cases?
That doesn't seem right!
I think one confounding factor might be testing.
```{r}
ggplot(recent,aes(x=gdp_per_capita,y=total_tests_per_thousand))+
  geom_point()+
  geom_smooth(method='lm')
summary(lm(gdp_per_capita~total_tests_per_thousand,recent))
```
```{r}
ggplot(recent,aes(x=total_cases_per_million,y=total_tests_per_thousand))+
  geom_point()+
  geom_smooth(method='lm')
summary(lm(total_cases_per_million~total_tests_per_thousand,recent))
```
Since there is high correlation between gdp and cases, it seems likely that the data doesn't tell the full story.

### Hypothesis test

```{r}
## A hypothesis test:
## Are no. of tests per thousand and no. of cases per million independent
## Let's do a chi squared test
### but first
frame2 = filter(recent,!(is.na(total_tests_per_thousand)))
ggplot(frame2,aes(x=total_cases_per_million,y=total_tests_per_thousand))+
  geom_point()+
  geom_smooth(method='lm')
summary(lm(total_cases_per_million~total_tests_per_thousand,frame2))
###
avg_tests = mean(frame2$total_tests_per_thousand)
avg_cases = mean(frame2$total_cases_per_million)
above_avg_test = filter(frame2,total_tests_per_thousand>avg_tests)
below_avg_test = filter(frame2,total_tests_per_thousand<avg_tests)
aa = length(filter(above_avg_test,total_cases_per_million>avg_cases)$iso_code)
ab = length(filter(above_avg_test,total_cases_per_million<avg_cases)$iso_code)
ba = length(filter(below_avg_test,total_cases_per_million>avg_cases)$iso_code)
bb = length(filter(below_avg_test,total_cases_per_million<avg_cases)$iso_code)
actual = c(aa,ab,ba,bb)
evaa = (aa+ab)*(aa+ba)/(aa+ab+ba+bb)
evab = (aa+ab)*(ab+bb)/(aa+ab+ba+bb)
evba = (ba+bb)*(ba+aa)/(aa+ab+ba+bb)
evbb = (bb+ab)*(bb+ba)/(aa+ab+ba+bb)
exp = c(evaa,evab,evba,evbb)
chisqr = sum(((actual-exp)^2)/exp)
pval = (1-pchisq(chisqr,df=1))*100
pval
```
Since the p value is 1.32%
We can see that these are clearly dependent
### Some Linear Modelling

```{r}

## okay so now linear modelling
### training
usa_data = filter(corona,iso_code=="USA")
usa_data$dayno = 1:147
ggplot(usa_data,aes(x=total_tests_per_thousand,y=total_cases_per_million))+
  geom_point()+
  geom_smooth(method='lm')
plot(total_cases_per_million~total_tests_per_thousand,usa_data)
fit1 = lm(total_cases_per_million~total_tests_per_thousand+I(total_tests_per_thousand^2)+I(total_tests_per_thousand^4),usa_data)
fit1
summary(fit1)
plot(total_cases_per_million~total_tests_per_thousand,usa_data)
x = usa_data$total_tests_per_thousand
points(x[!is.na(x)],fitted(fit1),col='blue',pch=8)
### model has been created

## now to see if it predicts well

## Inputting new data

corona_updated = read.csv("~/Downloads/owid-covid-data-updated.csv")
usa_new = filter(corona_updated,iso_code=="USA")
usa_new$dayno = 1:length(usa_new$total_cases_per_million)
##predict(fit1)
test_data = filter(usa_new,(dayno<158 & dayno>146))
prediction_data = predict(fit1,data.frame(total_tests_per_thousand=test_data$total_tests_per_thousand),interval="confidence")
predicted_values = prediction_data[1:11]
test_data$predicted_values = predicted_values
ggplot(test_data)+
  geom_point(aes(x=total_tests_per_thousand,y=total_cases_per_million,col='red'))+
  geom_point(aes(x=total_tests_per_thousand,y=predicted_values,col='blue'))
```